Best Practices for Multiple Alchemical Free Energy Calculations [Article v\versionnumber]

### 2025-04-09 paper/manuscript.tex
adding the connnecting section to the next bit.
Refined text.

>>>>

General simulation planning can be conceptualized as a three-component framework. Starting, the process involves identifying all mappings of the involved molecules (e.g. atom mappings), which define the structural relationships and transformations between molecule A and molecule B. Subsequently, these changes, along with their shared structural elements, are evaluated by estimating the difficulty of the transformations usually with predefined scoring functions. The final step entails constructing the perturbation graph using the scores from the prior step, which serves as a fundamental tool setting up a simulation campaign.

\paragraph{Atom mappings}
For minor molecular transformations, like in most large-scale approaches, single-coordinate set approaches are typically utilized due to their relatively low computational cost. These methods necessitate atom mappings to ensure their effective application. In large-scale implementations, the efficiency and robustness of automated mapping techniques are critical factors in these endavors.

%SIS
Subgraph isomorphism solver (SIS)-based maximum common substructure (MCSS) search methods currently represent the most widely used approach for addressing atom mapping challenges, at the current timepoint. Several cheminformatics tools implementing these solutions are readily available, including pmx~\cite{gapsys2015pmx}, LOMAP~\cite{liu2013lead}, FESetup~\cite{loeffler2015fesetup}, ProtoCaller~\cite{suruzhon2020protocaller}, and SMArt~\cite{petrov2021perturbation}.

These methods exhibit high accuracy in processing topological 2D information. However, hybrid approaches such as pmx~\cite{gapsys2015pmx} and LOMAP~\cite{liu2013lead}, which incorporate 3D structural data, enhance robustness in stereochemical considerations and molecular alignment, requiring well-prepared input coordinates. Regarding computational efficiency, SIS solvers must contend with the inherent complexity of solving an NP-complete problem~\cite{raymond2002maximum, Duesbury2018Comparison}. The use of an initial MCSS seed can mitigate this computational burden~\cite{raymond2002maximum}.

%Geometrz
An alternative to SIS-based MCSS methods is geometric bipartite matching for atom mapping, as implemented in tools such as pmx~\cite{gapsys2015pmx} and Kartograf~\cite{ries2024kartograf}. These algorithms enable rapid mappings due to their reduced computational complexity ($O(N) = N^3$), provided that well-aligned input geometries are available~\cite{ries2024kartograf}, through e.g. restraint-docking or prior-alignment. To accommodate topological variations, Kartograf applies a subsequent filtering step following the 3D mapping process.

% Short discussion of robustness and speed
Both mapping approaches present distinct advantages and limitations, each associated with specific prerequisites. Optimizing the robustness, and ignoring computational effort (e.g. mitigated by parallelization), it may be beneficial to employ both methods and implement a scoring mechanism to evaluate their performance. Most mappings are expected to be highly similar, this strategy could enhance robustness in the mapping process.

In terms of computational efficiency, the choice between methods largely depends on input conditions. If molecular structures are pre-aligned, the geometric approach offers a substantial speed advantage. Conversely, when ligands lack predefined 3D coordinates, the SIS-based method provides a easy solution, allowing coordinate generation at a later stage. This nuanced selection process ensures optimal performance based on the specific characteristics of the molecular system under investigation.

In combination with a suitable mapping scoring function the next steps is to build a pertubation map.
